Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-11 paper/paper.tex
edited bayesian section

Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge. Solving the storage and compute constraints encourages development
of model compress and sparsification methods, that focus on favourable tradeoff between
performance and size. Most popular techniquees include weight pruning \cite{citation_needed},
model quantization from IEEE754 floating point to integer-based arithmetic \cite{citation_needed},
and dropout. The key assumption is that the more explicitly zero parameters, the less
computations is needed.

@@

The use $\cplx$-valued networks in applications is not new: earliest mentions include
\cite{citation_needed}. More recently they were applied to music transcription and speech
recognition \cite{trabelsi_deep_2017,yang_complex_2019}. The key motivation behind adoption
of $\cplx$-networks is the fact that complex numbers are inherent in digital signal processing
\cite{citation_needed}, acoustic and radio signal analysis \cite{citation_needed}, and
quantum systems modelling \cite{citation_needed} etc. All these applications require
compliance with algebraic operations in field of complex numbers.

@@

The posterior itself is computationally intractable, save for the relatively simple cases
\cite{citation_needed}, and therefore exact inference is traded for tractability and scalability
of Variational Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach
recasts posterior inference into variational optimization problem
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its member have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \cite{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \cite{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

Kullback-Leibler divergence of density $q$ from $p$ is
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
and is the natural choice for $\rho$, the main reason being that it satisfies the following
identity for well-behaved $q$ and $p$:
\begin{equation}  \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This key identity yield an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta, \phi, \lambda}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is to use EM-algorithm.
Other variational objectives are possible, provided $p(\omega \mid D)$ is probed only
through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \cite{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \cite{kingma_auto-encoding_2014},
is an differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be also
replaced by differentiable unbiased Monte-Carlo estimators, \cite{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}

@@

For a particular family of posterior approximation, variational inference can be used
as a regularization method \cite{kingma_variational_2015} and as model sparsification
technique \cite{molchanov_variational_2017}. \cite{kingma_variational_2015} propose
Variational Dropout, by studying Binary dropout, DropConnect and Gaussian dropout,
\cite{hinton_improving_2012,wan_regularization_2013}, \cite{srivastava_dropout_2014},
and \cite{wang_fast_2013} through the lens of variational inference methods. The study
argues that the multiplicative noise, introduced by dropout, induces a distribution
equivalent to a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $\xi_j \sim p(\xi_j)$
iid from some $p_\xi$. Binary dropout with rate $p \in (0, 1)$ uses $
  p_\xi
    = \mathcal{B}\bigl(
      \{0, \tfrac1{1-p}\}, 1-p
    \bigr)
$, whereas in Gaussian dropout $
  \xi \sim \mathcal{N}(1, \alpha)
$ for $\alpha = \tfrac{p}{1-p}$. This suggests making $\alpha$ a variational parameter
and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
\cite{kingma_variational_2015} propose \textit{Variational Dropout} with mean field
Gaussian approximation $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. This unravels the divergence term in \eqref{eq:elbo} into a sum of the following terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
In general, there is tractable expression for log-moment of a non-central $\chi^2_m$ random
variable with odd $m$, that sits under the expectation in \eqref{eq:expect-improper-term-cplx}.
To circumvent this, \cite{kingma_variational_2015} approximate the expectation by a non-linear
polynomial regression for truncated $\alpha \in (0, 1)$. This approximation was refined in
\cite{molchanov_variational_2017}, both in terms of the domain of $\alpha$ and accuracy. The
paper considered regression with sigmoid and soft-plus basis functions fit to the Monte-Carlo
estimator of \eqref{eq:improper-kl-div-real} for $\alpha$ varying over a fine logarithmic grid.
In sec.~\ref{sub:real-chisq-grad} we verify the approximation and confirm its accuracy, by
comparing its derivative to the exact derivative of \eqref{eq:improper-kl-div-real} with
respect to $\log \alpha$.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Flexibility in dependence of $q_\theta$ on $\alpha$ allows structured dropout with shared
$\alpha$ across groups of parameters $\omega$. \cite{molchanov_variational_2017} study
Variational Dropout from a model sparsification perspective, optimizing $\alpha$ for each
individual parameter. Another variant of Bayesian dropout, called \textit{Automatic Relevance
Determination}, is developed in \cite{kharitonov_variational_2018}. The key difference is
that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\cite{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\lambda$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

Another contribution of \cite{kingma_variational_2015} is the analysis of the \textit{local
reparameterization trick}, proposed in \cite{wang_fast_2013} to speed up Gaussian dropout,
as a gradient variance reduction technique for \eqref{eq:elbo}. This trick utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
The stochastic output of a linear layer $
  y = W^\top x + b
$ with $
  W \in \mathbb{R}^{n\times m}
$ and independent $
  W_{ij} \sim \mathcal{N}(\mu_{ij}, \alpha_{ij} \mu_{ij}^2)
$ is equivalent in distribution to
\begin{equation}  \label{eq:r-gauss-trick}
    y \sim \prod_i \mathcal{N}\Bigl(
          \sum_j \mu_{ij} x_j + b_i,
          \sum_j \alpha_{ij} \mu_{ij}^2 x_j^2
      \Bigr)
    \,.
\end{equation}
%
\cite{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient. \cite{molchanov_variational_2017} propose an improvement by introducing
\textit{additive noise parameterization}, which is relevant to model sparsification.
The paper argues that $(\mu, \alpha)$ parameterization should be abandoned in favour
of an equivalent $(\mu, \sigma^2)$ with the score $\alpha$ calculated using $
  \tfrac{\sigma^2}{\mu^2}
$ when needed. This makes $
  \nabla_\mu \tilde{\mathcal{L}}
$, which is shown in \cite{molchanov_variational_2017} to be independent of the noise,
injected by the trick \eqref{eq:r-gauss-trick}, and further stabilizes gradients in SGVB
\eqref{eq:elbo}.

>>>>

Deep neural networks are an integral part of machine learning and data science toolset
for practical data-driven problem solving. Despite recent advances in generalization
and optimization theory specific to deep networks, deploying in actual embedded hardware
remains a challenge. Solving the storage and compute constraints encourages development
of model compression and sparsification methods, that focus on favourable tradeoff between
performance and size. Most popular techniques include weight pruning \cite{citation_needed},
model quantization from IEEE754 floating point to integer-based arithmetic \cite{citation_needed},
and dropout. The key assumption is that the more explicitly zero parameters there is,
the less computations is needed.

@@

Application of complex-valued networks is not new, \cite{citation_needed}. More recently
they were applied to music transcription and speech recognition \cite{trabelsi_deep_2017},
and \cite{yang_complex_2019}. The key motivation behind adoption of $\cplx$-networks is
pervasiveness of $\cplx$-valued data in digital signal processing \cite{citation_needed},
acoustic and radio signal analysis \cite{citation_needed}, and quantum systems modelling
\cite{citation_needed}, etc. All these applications require compliance with algebraic
operations in the field of complex numbers.

@@

The posterior itself is computationally intractable, save for the relatively simple cases
\cite{citation_needed}, and therefore exact inference is traded for tractability and scalability
of Variational Inference. Instead of deriving $p(\omega \mid D)$ using the rule, this approach
recasts posterior inference into variational optimization problem: finding $q$ that is close
to $p(\omega \mid D)$ in terms of a proximity score $\rho$
\begin{equation}  \label{eq:variational-progam}
  q_*
    \in \arg \min_{\theta} \rho\bigl(
      q_\theta(\omega), p(\omega \mid D)
    \bigr)
    \,,
\end{equation}
over a tractable parametric family of distributions on the parameter space $
  \mathcal{Q} = \{Q_\theta(d\omega) \colon \theta \in \Theta\}
$, $\theta$ -- generic variational parameter. The class $\mathcal{Q}$ is picked so that
its members have tractable densities $
  q_\theta(\omega) d\omega
$, can be sampled from and possess tractable $\log$-derivatives $
  \omega \mapsto \nabla_\theta \log q_\theta(\omega)
$, \cite{williams_simple_1992}, or can be represented as differentiable transformations
of non-parametric noise, \cite{kingma_auto-encoding_2014,figurnov_implicit_2019}.
% (reparameterization trick, pathwise gradieent) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1

Kullback-Leibler divergence of density $q$ from $p$ is
\begin{equation}  \label{eq:kl-div-def}
  KL(q \| p)
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q}
      \log{q(\omega)}
    - \mathbb{E}_{\omega \sim q}
      \log{p(\omega)}
    \,,
\end{equation}
% and, despite it not being a metric and requiring distributions with nested support,
and is the natural choice for $\rho$, the main reason being that it satisfies the following
identity for well-behaved densities:
\begin{equation}  \label{eq:kl-div-master}
  \log p(D)
    % = \mathbb{E}_{q(\omega)} \log \frac{p(D, \omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log \frac{
    %   p(D \mid \omega) \pi(\omega)
    % }{q(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    % = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    % - \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{\pi(\omega)}
    % + \mathbb{E}_{q(\omega)} \log \frac{q(\omega)}{p(\omega \mid D)}
    = \mathbb{E}_{q(\omega)} \log{p(D \mid \omega)}
    - KL(q(\omega) \| \pi(\omega))
    + KL(q(\omega) \| p(\omega \mid D))
    \,.
\end{equation}
This identity yields an equivalent problem for \eqref{eq:variational-progam}:
maximizing the \textit{evidence lower bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
\begin{aligned}
  & \underset{\theta, \phi, \lambda}{\text{maximize}}
    & \mathcal{L}(\theta; \phi, \lambda)
      & = \mathbb{E}_{\omega \sim q_{\theta}}
          \log p_{\phi}(D \mid \omega)
        - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
        % \mathbb{E}_{\omega \sim q_{\theta}}
          % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % \\
  % & &
  %     & = \mathbb{E}_{\omega \sim q_{\theta}}
  %         \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %       + \mathbb{H}(q_{\theta})
  \,,
\end{aligned}
\end{equation}
where the prior and likelihood are allowed to depend on explicit parameters. The most
commonly used algorithm to optimize \eqref{eq:elbo_general} is EM-algorithm. Other
variational objectives are possible, provided $p(\omega \mid D)$ is probed only through
$\log p(D \mid \omega)$ and $\log \pi(\omega)$, \cite{ranganath_operator_2018}.

\textit{Stochastic Gradient Variational Bayes} (SGVB), proposed by \cite{kingma_auto-encoding_2014},
is a differentiable unbiased Monte-Carlo estimator of \eqref{eq:elbo_general}, which makes
it possible to employ stochastic gradient-based optimization methods. If $
  \omega \sim q_{\theta}
$ is equivalent in distribution to $
  \omega = g(\varepsilon; \theta)
$ for some non-parametric random variable $
    \varepsilon \sim p_\varepsilon
$ and differentiable $
  g(\varepsilon; \theta)
$, then the SGVB is
\begin{equation}  \label{eq:sgvb}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(z_i \mid \omega_i)b
      + \frac1{L} \sum_{l=1}^L
        \log p_{\phi}(D \mid \omega_{il})
          \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta}(\omega) \| \pi(\omega))
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{z,\omega \sim B}
          % \log p(z \mid \omega)
    \,, 
\end{equation}
where $
  (\varepsilon_{l})_{l=1}^L
$ is an iid sample from $p_\varepsilon$. Intractable KL-divergence terms can be replaced
by differentiable unbiased Monte-Carlo estimators, \cite{kingma_auto-encoding_2014}.
%
For the purposes of this study, we assume that the observed data is independent, and
use minibatch version of \eqref{eq:sgvb} with $L=1$:
\begin{equation}  \label{eq:elbo}
  \widetilde{\mathcal{L}}(\theta; \phi, \lambda)
    = - KL(q_{\theta}(\omega) \| \pi_{\lambda}(\omega))
      + \frac{N}{M} \sum_{i=1}^M
        \log p_{\phi}(z_i \mid g(\varepsilon_i; \theta))
    \,,
\end{equation}
for a random subsample $(z_i)_{i=1}^M$ from $D$ and iid $
  (\varepsilon_i)_{i=1}^M \sim p_\varepsilon
$.
%
\todo{DSVI}

@@

With special family of posterior approximation, variational inference can be used as
a regularization method, \cite{kingma_variational_2015}, and as model sparsification
technique, \cite{molchanov_variational_2017}. The paper \cite{kingma_variational_2015}
considers Dropout \cite{hinton_improving_2012}, DropConnect, \cite{wan_regularization_2013},
and Gaussian dropout, \cite{srivastava_dropout_2014,wang_fast_2013} through the lens
of Bayesian inference methods and proposes \textit{Variational Dropout}. It argues that
the multiplicative noise introduced by these methods induces a distribution equivalent to
a fully factorized variational posterior of the form $
  q_\theta(\omega) = \prod_j q_{\theta}(\omega_j)
$, where $q_{\theta}(\omega_j)$ is $\omega_j = \mu_j \xi_j$ with $
  \xi_j \sim p_\theta(\xi_j)
$ iid from some $p_\theta(\xi)$.
%
% Binary dropout with rate $p \in (0, 1)$ uses $
%   p_\xi
%     = \mathcal{B}\bigl(
%       \{0, \tfrac1{1-p}\}, 1-p
%     \bigr)
% $, whereas in Gaussian dropout $
%   \xi \sim \mathcal{N}(1, \alpha)
% $ for $\alpha = \tfrac{p}{1-p}$.
% This suggests making $\alpha$ a variational parameter
% and optimizing it in \eqref{eq:elbo_general} with an appropriate penalty term. Thus,
%
Variational Dropout with assumes mean field Gaussian approximation $
  q_\theta(\omega)
    = \prod_j \mathcal{N}(\mu_j, \alpha_j \mu_j^2)
$ and factorized log-uniform prior $\pi(\omega)$ with $
  \pi(\omega_j) \propto \lvert \omega_j \rvert^{-1}
$. This unravels the divergence term in \eqref{eq:elbo} into a sum of the following
analytically intractable terms:
\begin{equation}  \label{eq:improper-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    \propto
      \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
        \log{\Bigl\lvert \tfrac1{\sqrt{\alpha}} + \varepsilon \Bigr\rvert^2}
      % \frac12 \mathbb{E}_{\varepsilon \sim \mathcal{N}(0, 1)}
      %   \log{\bigl\lvert 1 + \sqrt{\alpha} \varepsilon \bigr\rvert^2}
      % - \frac12 \log{\alpha}
  \,.
\end{equation}
\cite{kingma_variational_2015} approximate this expression by a non-linear polynomial
regression for \verify{truncated} $\alpha \in (0, 1)$. This approximation was refined in
\cite{molchanov_variational_2017}, both in terms of the domain of $\alpha$ and accuracy. The
paper considered regression with sigmoid and soft-plus basis functions fit to the Monte-Carlo
estimator of \eqref{eq:improper-kl-div-real} for $\alpha$ varying over a fine logarithmic grid.
In sec.~\ref{sub:real-chisq-grad} we verify the approximation and confirm its accuracy, by
comparing its derivative to the exact derivative of \eqref{eq:improper-kl-div-real} with
respect to $\log \alpha$.
% (!) good for SGVB: gradient-based methods require unbiased gradient estimates

Flexibility in specification of $q_\theta$ enables structured dropout with shared
$\alpha$ across groups of parameters $\omega$. \cite{molchanov_variational_2017} study
Variational Dropout from a model sparsification perspective, optimizing $\alpha$ for each
individual parameter. Another variant of Bayesian dropout, called \textit{Automatic Relevance
Determination}, is developed in \cite{kharitonov_variational_2018}. The key difference is
that the log-uniform factor $\pi(\omega_{ij})$ is replaced by a proper Gaussian prior $
  \mathcal{N}(0, \tau^{-1}_{ij})
$ with learnable precision $\tau_{ij} > 0$. This method, known as \textit{Empirical Bayes}
\cite{citation_needed}, fits a prior distribution while performing variational inference.
Maximizing \eqref{eq:elbo} over $\lambda$, holding other parameters fixed, yields $
  \tau^* = {(\mu^2 + \sigma^2)}^{-1}
$, whence
\begin{equation}  \label{eq:ard-kl-div-real}
  KL(q_\theta(\omega) \| \pi(\omega))
    = \log{\bigl(1 + \tfrac1{\alpha} \bigr)}
    \,.
\end{equation}

Another contribution of \cite{kingma_variational_2015} is the analysis of the effects of the
\textit{local reparameterization trick} on the variance of the gradient of \eqref{eq:elbo}.
This trick, proposed in \cite{wang_fast_2013} to speed up Gaussian dropout, utilizes the
closure of Gaussians under affine transformations, thereby translating uncertainty from the
global parameter noise to local noise in intermediate outputs within the network.
%
The stochastic output of a linear layer $
  y = W^\top x + b
$ with $
  W \in \mathbb{R}^{n\times m}
$ and independent $
  W_{ij} \sim \mathcal{N}(\mu_{ij}, \alpha_{ij} \mu_{ij}^2)
$ is equivalent in distribution to
\begin{equation}  \label{eq:r-gauss-trick}
    y \sim \prod_i \mathcal{N}\Bigl(
          \sum_j \mu_{ij} x_j + b_i,
          \sum_j \alpha_{ij} \mu_{ij}^2 x_j^2
      \Bigr)
    \,.
\end{equation}
%
\cite{kingma_variational_2015} show, that for the case of fully factorized Gaussian
posterior approximations, the trick decorrelates ELBO estimates within the minibatch,
making the gradient estimator $\nabla_\theta \tilde{\mathcal{L}}$ more statistically
efficient. \cite{molchanov_variational_2017} propose an improvement by introducing
\textit{additive noise parameterization}, which is relevant to model sparsification.
The paper argues that $(\mu, \alpha)$ parameterization should be abandoned in favour
of an equivalent $(\mu, \sigma^2)$ with the score $\alpha$ calculated using $
  \tfrac{\sigma^2}{\mu^2}
$ when needed. This makes $
  \nabla_\mu \tilde{\mathcal{L}}
$, which is shown in \cite{molchanov_variational_2017} to be independent of the noise,
injected by the trick \eqref{eq:r-gauss-trick}, and further stabilizes gradients in SGVB
\eqref{eq:elbo}.
